Using structural and evolutionary information to correct RNA pyrosequencing errors in non-coding RNAs

### 2013-05-25 JCB/methods_JCB.tex
modifs figs

To that purpose, one leverages the \emph{Inside-Outside} construction. Namely, while computing the total contribution of sequences featuring a base $a'\in\B$ at position $i$, one must consider the three following situations:
\begin{itemize}
\item {\bf Case 1: Unpaired position.} In this case, the supporting sequences are simply those excluding the structure $\ub$ at position $i$, where a base $c\in\B$ was formerly found. Summing over any possible number of mutations for the \emph{outside} region, one obtains:
\begin{equation}
 \mathcal{W}^M_{\substack{i, [a]}} =  \sum_{m=\Kron_{a,c}}^{M}
			\Y{\ub}{m-\Kron_{a,c}}{a,a}.
\label{eq:combine_A}
\end{equation}
\item {\bf Case 2: Position is paired rightward.} Here, position $i$ is paired to some position $k>i$, whose content impacts the base-pair contribution. Any sequence having $a$ at position $i$ can be decomposed as a base-pair, nesting a sequence for the structure $\Struct'$ on the interval $[i+1,k-1]$, and an outside sequence, excluding the structure $\op\Struct'\cp$.
Consequently, let $c$ and $d$ be the original nucleotides at positions $i$ and $k$, one has:
\begin{equation}
 \mathcal{W}^M_{\substack{i, [a]}} =  
			\sum_{m=0}^{M}
			\sum_{\substack{b\in \B\\\Kron_{ab,cd}\leq m}}
			\sum_{m'=0}^{m-\Kron_{ab,cd}}
     	 e^{\frac{-\EBP{i,k}{\varnothing}{ab}}{RT}}
			\cdot\Y{\op\Struct'\cp}{m-\Kron_{ab,cd}}{a,b}
			\cdot\Z{\Struct'}{m'}{a,b}
\label{eq:combine_B}
\end{equation}
\item {\bf Case 3: Position is paired leftward.} This case is symmetrical to the previous one, with the exception that $k<i$.
Consequently, let $c$ and $d$ be the original nucleotides at positions $i$ and $k$, one has:
\begin{equation}
 \mathcal{W}^M_{\substack{i, [a]}} =  
			\sum_{m=0}^{M}
			\sum_{\substack{b\in \B\\\Kron_{ba,cd}\leq m}}
			\sum_{m'=0}^{m-\Kron_{ba,cd}}
     	 e^{\frac{-\EBP{k,i}{\varnothing}{ba}}{RT}}
			\cdot\Y{\op\Struct'\cp}{m-\Kron_{ba,cd}}{b,a}
			\cdot\Z{\Struct'}{m'}{b,a}
\label{eq:combine_C}
\end{equation}
\end{itemize}

>>>>

\begin{figure}[t]\centering
\includegraphics[scale=\ScaleDP]{FigDPCombineWrapper}
\caption{Combining the inside and outside contributions to compute the total Boltzmann weight of all sequences having a given base (case 1) or base-pair (cases 2 and 3) at a given position.\label{fig:combine}}
\end{figure}

To that purpose, one leverages the \emph{Inside-Outside} construction, as  illustrated by Figure~\ref{fig:combine}. Namely, while computing the total contribution of sequences featuring a base $a'\in\B$ at position $i$, one must consider the three following situations:
\begin{itemize}
\item {\bf Case 1: Unpaired position.} In this case, the supporting sequences are simply those excluding the structure $\ub$ at position $i$, where a base $c\in\B$ was formerly found. Summing over any possible number of mutations for the \emph{outside} region, one obtains:
\begin{equation}
 \mathcal{W}^M_{\substack{i, [a]}} =  \sum_{m=\Kron_{a,c}}^{M}
			\Y{\ub}{m-\Kron_{a,c}}{a,a}.
\label{eq:combine_A}
\end{equation}
\item {\bf Case 2: Position is paired rightward.} Here, position $i$ is paired to some position $k>i$, whose content impacts the base-pair contribution. Any sequence having $a$ at position $i$ can be decomposed as a base-pair, nesting a sequence for the structure $\Struct'$ on the interval $[i+1,k-1]$, and an outside sequence, excluding the structure $\op\Struct'\cp$.
Consequently, let $c$ and $d$ be the original nucleotides at positions $i$ and $k$, one has:
\begin{equation}
 \mathcal{W}^M_{\substack{i, [a]}} =  
			\sum_{m=0}^{M}
			\sum_{\substack{b\in \B\\\Kron_{ab,cd}\leq m}}
			\sum_{m'=0}^{m-\Kron_{ab,cd}}
     	 e^{\frac{-\EBP{i,k}{\varnothing}{ab}}{RT}}
			\cdot\Y{\op\Struct'\cp}{m-\Kron_{ab,cd}}{a,b}
			\cdot\Z{\Struct'}{m'}{a,b}
\label{eq:combine_B}
\end{equation}
\item {\bf Case 3: Position is paired leftward.} This case is symmetrical to the previous one, with the exception that $k<i$.
Consequently, let $c$ and $d$ be the original nucleotides at positions $i$ and $k$, one has:
\begin{equation}
 \mathcal{W}^M_{\substack{i, [a]}} =  
			\sum_{m=0}^{M}
			\sum_{\substack{b\in \B\\\Kron_{ba,cd}\leq m}}
			\sum_{m'=0}^{m-\Kron_{ba,cd}}
     	 e^{\frac{-\EBP{k,i}{\varnothing}{ba}}{RT}}
			\cdot\Y{\op\Struct'\cp}{m-\Kron_{ba,cd}}{b,a}
			\cdot\Z{\Struct'}{m'}{b,a}
\label{eq:combine_C}
\end{equation}
\end{itemize}
